Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-16 curves-write-up/arXiv versions/spin-cring-arXiv-v2.tex
add colon

Let $X$ be a smooth proper geometrically-integral algebraic curve of genus $g$ over a field $\Bk$.
It is well known that the canonical sheaf $\Omega _X,$ with
associated canonical divisor $K_X$, determines the {\bf canonical
map } $\pi: X \rightarrow \BP_\Bk^{g - 1}$. Then, the {\bf canonical
ring} is defined to be
\begin{align*}
	R(X, K_X) = \bigoplus_{d \geq 0} H^0(X, dK_X),
\end{align*}

@@

\begin{defn}
\label{defn:order-sup}
Let $D \in \di \sx.$ If $z$ is a rational section of $\sco(D)$ denote the order of
zero of $z$ at $P$ by $\ord_P^D(z)$.
\end{defn}

@@

\begin{lem}
\label{lem:relations_from_generators_induction} 
Suppose $\halfcan$, $\halfcan' \in \BQ$ $\otimes$ $\di X$ with $\halfcan=\halfcan'+\frac{\alpha}{\beta}P$, such that $R_{\halfcan'}$ generated by $x_1, \ldots, x_m$
and $R_{\halfcan}$ is minimally generated by $y_1, \ldots, y_n$ over 
$R_{\halfcan'}$.  Let $I'$ and $I$ be the ideals relations of $\phi':\Bk[x_1, \ldots, x_m]\to R_{\halfcan'}$ and $\phi:\Bk[x_1, \ldots, x_m, y_1, \ldots y_n]\to R_{\halfcan}$ respectively.
Suppose there are sets of monomials $S\subseteq R_\halfcan-R_{\halfcan'}$ and $T\subseteq R_\halfcan-(S\cup R_\halfcan')$, and a monomial ordering $\prec$ such 
\begin{enumerate}
\item $S$ forms a $\Bk$-basis for $R_\halfcan$ over $R_\halfcan'$
\item $T \succ S\succ \Bk[x_1, \ldots, x_n]$ {\rm(}meaning all monomials in $T$ are bigger than all monomials in $S$ which are bigger than all monomials in $\Bk[x_1, \ldots, x_n]${\rm)}
\item  All monomials in $\Bk[x_1, \ldots, x_m, y_1, \ldots, y_n]$ lie in 
\begin{align*}
	S \; \cup\; \langle T\rangle \; \cup \; \initial(I') \Bk[x_1, \ldots, x_m, y_1, \ldots, y_n] \; \cup \; \Bk[x_1, \ldots, x_m]
\end{align*}
\end{enumerate}
Then,
\begin{align*}
	\initial(I) & = \initial(I') \Bk[x_1, \ldots, x_m, y_1, \ldots, y_n]
	+ \langle T \rangle.
\end{align*}
\end{lem}

@@

\item[(b)] Choose an ordering $\prec$ on $\Bk[u, x_1, \ldots, x_m]$ such that
	\[
		\ord_u(f) < \ord_u(h) \implies f\prec h.
	\]
	Equip $\Bk[y_1, \ldots, y_n]$ with graded $P$-lex, as defined in
	Definition ~\ref{defn:graded-p-lex}, and equip $\Bk[y_1, \ldots,
	y_n] \otimes \Bk[u, x_1, \ldots, x_m]$ with block order.
	If $I'$ is the ideals of relations of $\Bk[u, x_1, \ldots, x_m]
	\to R_{\halfcan'}$ and $I$ is the ideal of relations of
	$\Bk[u, x_1, \ldots, x_m, y_1, \ldots, y_n]	\to R_{\halfcan}$, then

>>>>

Let $X$ be a smooth proper geometrically-connected algebraic curve of genus $g$ over a field $\Bk$.
It is well known that the canonical sheaf $\Omega _X,$ with
associated canonical divisor $K_X$, determines the {\bf canonical
map } $\pi: X \rightarrow \BP_\Bk^{g - 1}$. Then, the {\bf canonical
ring} is defined to be
\begin{align*}
	R(X, K_X) := \bigoplus_{d \geq 0} H^0(X, dK_X),
\end{align*}

@@

\begin{defn}
\label{defn:order-sup}
Let $D \in \di \sx.$ If $z \neq 0$ is a rational section of $\sco(D)$ denote the order of
zero of $z$ at $P$ by $\ord_P^D(z)$.
\end{defn}

@@

\begin{lem}
\label{lem:relations_from_generators_induction} 
Suppose $\halfcan$, $\halfcan' \in \BQ$ $\otimes$ $\di X$ with $\halfcan=\halfcan'+\frac{\alpha}{\beta}P$, such that $R_{\halfcan'}$ generated by $x_1, \ldots, x_m$
and $R_{\halfcan}$ is minimally generated by $y_1, \ldots, y_n$ over 
$R_{\halfcan'}$.  Let $I'$ and $I$ be the ideals of relations of $\phi':\Bk[x_1, \ldots, x_m]\to R_{\halfcan'}$ and $\phi:\Bk[x_1, \ldots, x_m, y_1, \ldots y_n]\to R_{\halfcan}$ respectively.
Suppose there are sets of monomials $S\subseteq R_\halfcan-R_{\halfcan'}$ and $T\subseteq R_\halfcan-(S\cup R_\halfcan')$, and a monomial ordering $\prec$ such 
\begin{enumerate}
\item $S$ forms a $\Bk$-basis for $R_\halfcan$ over $R_\halfcan'$
\item $T \succ S\succ \Bk[x_1, \ldots, x_n]$ {\rm(}meaning all monomials in $T$ are bigger than all monomials in $S$ which are bigger than all monomials in $\Bk[x_1, \ldots, x_n]${\rm)}
\item  All monomials in $\Bk[x_1, \ldots, x_m, y_1, \ldots, y_n]$ lie in 
\begin{align*}
	S \; \cup\; \langle T\rangle \; \cup \; \initial(I') \Bk[x_1, \ldots, x_m, y_1, \ldots, y_n] \; \cup \; \Bk[x_1, \ldots, x_m]
\end{align*}
\end{enumerate}
Then,
\begin{align*}
	\initial(I) & = \initial(I') \Bk[x_1, \ldots, x_m, y_1, \ldots, y_n]
	+ \langle T \rangle.
\end{align*}
\end{lem}

@@

\item[(b)] Choose an ordering $\prec$ on $\Bk[u, x_1, \ldots, x_m]$ such that
	\[
		\ord_u(f) < \ord_u(h) \implies f\prec h.
	\]
	Equip $\Bk[y_1, \ldots, y_n]$ with graded $P$-lex, as defined in
	Definition ~\ref{defn:graded-p-lex}, and equip $\Bk[y_1, \ldots,
	y_n] \otimes \Bk[u, x_1, \ldots, x_m]$ with block order.
	If $I'$ is the ideal of relations of $\Bk[u, x_1, \ldots, x_m]
	\to R_{\halfcan'}$ and $I$ is the ideal of relations of
	$\Bk[u, x_1, \ldots, x_m, y_1, \ldots, y_n]	\to R_{\halfcan}$, then

### 2015-07-16 surfaces-write-up/surface-write-up.tex

The work in this paper also has relations to certain rings of Hilbert 
modular forms and Siegel modular forms. Recall that Hilbert and Siegel 
modular forms are both to higher dimensions variables. In the case of 
Hilbert Modular forms, one considers modular forms on an $m$ fold product 
of upper half planes, which transform under the action of finite index 
subgroups of the groups of positive determinant matrices. Similarly, Siegel 
modular forms are modular forms on the Siegel upper half space which are 
stable under the action of the symplectic group.

Voight and Zureick-Brown \cite[Proposition 6.1.5]{vzb:stacky}, which allows one to translate the analytic category of 1-dimensional complex orbifolds to the algebraic category of stacky curves. 
We expect that  will similarly allow one to generalize yield an eqivalence of categories between an appropriate subcategory of $m$-dimensional orbifolds and a subcategory of $m$-dimensional algebraic spaces.
In particular, this will allow us to translate rings of Hilbert Modular forms and Siegel modular forms whose underlying space is isomorphic to $\bp^m$ or a Hirzebruch surface to canonical rings on the corresponding algebraic surfaces. Thereby, our techniques would give a bound on generators and relations for Hilbert and Siegel modular forms with certain underlying spaces.
\todo{include references}

Another closely related topic is the Hassett-Kell program 
initiated in \todo{ was it actually initiated here? check vzb.} \cite{hassett:classical-and-minimal-models}, which aims to describe log canonical models of the form
\begin{align*}
	\overline {\mathscr M}_g(\alpha) := \oplus_{d \geq 0}H^0 \left( \overline {\mathscr M}_g, \lfloor d K_{\overline{\mathscr M}_g} + \alpha\delta \rfloor  \right) 
\end{align*}
in terms of certain moduli spaces. Fractional divisors on surfaces also naturally appear when considering the canonical ring of surfaces of kodaira dimension 1. A surface has Kodaira dimension 1 if and only if it is an elliptic surface, and their canonical ring are most naturally described in the setting of fractional divisors. \todo{Add references}

@@

\begin{rem}
\label{rem:proj-veronese}
For any divisor $D = \sum_{i = 0}^{n} \alpha_i D_i$ on $\bp^m$
with each $D_i$ a hypersurface, we can use the Veronese mappings
to view $D$ as a divisor on $\bp^{m'}$ for some $m' \in \bz_{\geq 0}$
with $D = \sum_{i = 0}^{n} \alpha_i H_i$ where $H_i$ are hypersurfaces
in $\bp^{m'}$.
\end{rem}

@@

\begin{rem}
\label{rem:proj-one-point-ind}
Note that the method of proof for Theorem ~\ref{thm:proj-one-point}
can be immediately generalized to work inductively. If $D = D' + H$
for some hyperplane $H$, then we can use the proof above to show
the canonical ring $R_D$ is generated over $R_D'$ by the same
generators and with the same relations given in Theorem
~\ref{thm:proj-one-point}.
\end{rem}

>>>>

The work in this paper also has relations to certain rings of Hilbert 
modular forms and Siegel modular forms. Recall that Hilbert and Siegel 
modular forms are both to higher dimensions variables. In the case of 
Hilbert Modular forms, one considers modular forms on a product of $m$
upper half planes, which transform under the action of finite index 
subgroups of the groups of positive determinant matrices. Similarly, Siegel 
modular forms are modular forms on the Siegel upper half space which are 
stable under the action of the symplectic group.

By a result by Voight and Zureick-Brown \cite[Proposition 6.1.5]{vzb:stacky}, one can translate the analytic category of 1-dimensional complex orbifolds to the algebraic category of stacky curves. 
We expect that  will similarly allow one to generalize yield an eqivalence of categories between an appropriate subcategory of $m$-dimensional orbifolds and a subcategory of $m$-dimensional algebraic spaces.
In particular, this will allow us to translate rings of Hilbert Modular forms and Siegel modular forms whose underlying space is isomorphic to $\bp^m$ or a Hirzebruch surface to canonical rings on the corresponding algebraic surfaces. Thereby, our techniques would give a bound on generators and relations for Hilbert and Siegel modular forms with certain underlying spaces.
\todo{include references}

Another closely related topic is the Hassett-Keel program 
initiated in \todo{ was it actually initiated here? check vzb.} \cite{hassett:classical-and-minimal-models}, which aims to describe log canonical models of the form
\begin{align*}
	\overline {\mathscr M}_g(\alpha) := \oplus_{d \geq 0}H^0 \left( \overline {\mathscr M}_g, \lfloor d K_{\overline{\mathscr M}_g} + \alpha\delta \rfloor  \right) 
\end{align*}
in terms of certain moduli spaces. Fractional divisors on surfaces also naturally appear when considering the canonical ring of surfaces of kodaira dimension 1. A surface has Kodaira dimension 1 if and only if it is an elliptic surface, and their canonical ring are most naturally described in the setting of fractional divisors. \todo{Add references}

Furthermore, this work extends the study of canonical rings of
$\bq$-divisors on curves, as done by O'Dorney \cite{dorney:canonical},
Voight--Zureick-Brown \cite{vzb:stacky}, and Landesman--Ruhm--Zhang
\cite{lrz:spin-cring}, to general projective space and Hirzebruch
surfaces.

\todo{more!}

@@

@@

\begin{rem}
\label{rem:proj-one-point-ind}
The method of proof for Theorem ~\ref{thm:proj-one-point}
can be immediately generalized to work inductively. If $D = D' + H$
for some hyperplane $H$, then we can use the proof above to show
the canonical ring $R_D$ is generated over $R_D'$ by the same
generators and with the same relations given in Theorem
~\ref{thm:proj-one-point}.
\end{rem}
